---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cbbAnalysis

<!-- badges: start -->
<!-- badges: end -->

The goal of cbbAnalysis is to scrape Sports Reference webpages to extract NCAA basketball statistics and stores them as tables. It enables analysis using tools such as random forest regression, allowing users to build samples of any size, dynamically predict matchups, and evaluate performance metrics.

## Installation

You can install the development version of cbbAnalysis like so:

``` r

install.packages("devtools")
devtools::install_github("HadenGwin/cbbAnalysis")

```

# Examples

Here is a basic example that demonstrates some of the functionality of the cbbAnalysis package:

## Scraping Matchups

```{r matchups-example}
library(cbbAnalysis)
# Get matchups for December 1, 2024
matchups <- matchupsDate("2024-12-01")
print(matchups)
```

## Calculating Four Factors Statistics

```{r four-factors}
boxscore_url <- "https://www.sports-reference.com/cbb/boxscores/2024-12-02-03-nevada.html"

four_factors <- create_four_factors(boxscore_url)
print(four_factors)
```

## Individual Team Statistics

``` {r team-stats}

team = "https://www.sports-reference.com/cbb/schools/indiana/2024.html"

teamStats <- calculate_team_metrics(team)
print(teamStats)
```

## train a model on a dynamic sample size

``` {r train-model}
start_date <- "2024-11-28"
end_date <- "2024-12-12"
num_games_per_day = 3
ntree = 699

newModel <- trainModel(start_date, end_date, num_games_per_day, ntree)
```


## predict matchups

``` {r predict-matchups}
team1_url <- "https://www.sports-reference.com/cbb/schools/indiana/2024.html"
team2_url <- "https://www.sports-reference.com/cbb/schools/connecticut/2024.html"

results <- trainModel(start_date = "2024-11-01", end_date = "2024-11-15", num_games_per_day = 5, ntree = 300)

predictions <- predict_matchup(team1_url, team2_url, results)
print(predictions)
```

